---
title: "Prediction on training quality"
author: "Dmitry Maskaev"
date: "Sunday, September 27, 2015"
output: html_document
---

##Overview
In this analysis we are going to explore data from Weight Lifting Exercise Dataset, build a prediction model and see the resuilts of its prediction. 

##Downloading dataset
First we download dataset to current folder.
```{r, eval=FALSE}
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",destfile = "train.csv",mode = "wb")
download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",destfile = "test.csv",mode = "wb")
```

##Reading data
Then we specify seed for reproducibility, load needed libraries and finally read data to dataframes.
```{r,cache=TRUE}
set.seed(1234)
library(randomForest)
library(caret)

train <- read.table(file = "train.csv",sep = ",",header = T,na.strings = c("NA","#DIV/0!"),check.names = T)
test <- read.table(file = "test.csv",sep = ",",header = T,na.strings = c("NA","#DIV/0!"),check.names = T)
```

##Data exploration
We make initial data exploration.
```{r}
head(train)
str(train)
```

##Data transformation
We remove first 5 columns as they won't be useful for prediction. As well we remove columns with NAs.
```{r,cache=TRUE}
train$X <- NULL
train$user_name <- NULL
train$raw_timestamp_part_1 <- NULL
train$raw_timestamp_part_2 <- NULL
train$cvtd_timestamp <- NULL
train <- train[,colnames(train)[!(colSums(is.na(train)) > 0)]]
```

Then we remove predictors with near-zero variance. 
```{r,cache=TRUE}
nzv <- nearZeroVar(train)
filteredtrain <- train[, -nzv]
```

And split training data to train and validation datasets.
```{r,cache=TRUE}
ind <- createDataPartition(y = filteredtrain$classe, list = F, p = 0.7)
train <- filteredtrain[ind,]
valid <- filteredtrain[-ind,]
```

##Building a model
After that we build a model. We'll use random forests for that with default options and Principal Component Analysis (PCA) pre-processing.
```{r,cache=TRUE,message=FALSE}
fit <- train(classe ~ ., data = train, method = "rf", preProcess = "pca")
```

##Prediction
Finally let's try make predictions with our model and see predictions results on the training set...
```{r}
pred <- predict(object = fit,newdata = train)
table(pred,train$classe)
```

...and on the validation set.
```{r}
pred2 <- predict(object = fit,newdata = valid)
table(pred2,valid$classe)
```
As we see the prediction with random forests is quite exact. But we are waiting a bit worse result on test data.
